Introducing ICBe: An Event Extraction Dataset from Narratives about International Crises

### 2023-11-20 replication_paper/psrm_draft/ICBe_manuscript.Rmd
AG rr 2 edits

A unique feature of the ontology is that thought, speech, and do behaviors can be nested into combinations, e.g. an offer for the U.S.S.R. to remove missiles from Cuba in exchange for the U.S. removing missiles from Turkey. Through compounding, the ontology can capture what players were said to have known, learned, or said about other specific fully described actions. Two other novel features of our data -- its re-programmamable coding and its inclusion of thought, speech, and action -- intersect to provide a unique contribution. No existing event data distinguishes thoughts, speeches, and actions and most are limited to just coding actions given they are the most readily observable and thus the full universe could at least theoretically be included.[^7] Yet, methods like archival work and expert interviews continue to improve identification of thoughts and speeches that help us understand the reasoning process underlying political decisions and what actually occurred behind the scenes. This is an improvement over the state of the art that omits these by design in allowing scholars to identify thoughts associated with things like crisis escalation as well as patterns of omission to be advanced in future work, but the modular nature of our coding means the inclusion of a non-random sample of thoughts and speeches is neither necessary nor does it have inevitable downstream effects on the rest of the coding in ways that produce unwanted bias.

@@

Our strategy for dealing with both problems is a plausibly objective yardstick called a synthetic historical narrative. We collect a large diverse corpus of narratives spanning timelines, encyclopedia entries, journal articles, news reports, websites, and government documents. Using natural language processing (fully described in SI Appendix 3.1), we identify details that appear across multiple accounts. The more accounts that mention a detail, the more central it is to understanding the true historical episode. The theoretical motivation is that authors face word limits which force them to pick and choose which details to include, and they choose details that serve the specific context of the document they are producing. With a sufficiently large and diverse corpus of documents, we can vary the context while holding the overall episode constant and see which details tend to be invariant to context. This presents a harder evaluation baseline than comparing ICBe's recall to just that of ICB since there are non-crisis aspects of these events that may be included in other narratives but are out of the scope of our data. For example, the nationalization of businesses in Cuba may be included as important context in the Cuban Missile Crisis in documents that do not focus on the crisis dimensions like ICB. Using this hard case, a recall measure of ICBe on the synthetic narratives thus serves as a way to evaluate the breadth of ICBe's ontology and potential application to non-crisis international events.

>>>>

A unique feature of the ontology is that thought, speech, and do behaviors can be nested into combinations, e.g. an offer for the U.S.S.R. to remove missiles from Cuba in exchange for the U.S. removing missiles from Turkey. Through compounding, the ontology can capture what players were said to have known, learned, or said about other specific fully described actions. Two other novel features of our data -- its re-programmamable coding and its inclusion of thought, speech, and action -- intersect to provide a unique contribution.

First, re-programmable coding means scholars can make modifications to the ontology we provide and re-run the code script to produce a modified dataset. For example, if thoughts and speech acts are immaterial for a particular research question, those higher-level categories can be modified in a way that does not produce inevitable downstream effects on the rest of the coding. This is necessary because unlike existing datasets where variable values are largely independent, the nested nature of the ICBe variables means filtering out speech acts also requires filtered out variables about the actors, locations, and timing of those speech acts.

Second, no existing event data distinguishes thoughts, speeches, and actions. In fact, most only try to code actions and entirely omit thoughts and speech acts despite recognition of their importance in international politics [@smithInternationalCrisesDomestic1998]. Scholars have opted against coding thoughts and speech acts because of a lack of confidence the full universe could be readily observed and consequently at least theoretically be included.[^7] But the perfect should not be the enemy of the good, and measurement challenges are only overcome after an initial attempt to estimate difficult to observe concepts of interest. The ICB narratives are one of the better sources for this endeavor due to the consistent use of high-quality primary source material that takes advantage of qualitative methods well-suited to identifying thoughts and speech acts like archival work and expert interviews. Important questions about potential bias remain like whether a thought was known at the time by those involved or only recorded long afterwards by historians, but data validation required there to be data to be validated. By design, existing datasets largely omit thoughts and speech acts and also treat the resulting dataset as the static product of hardwired coding decisions. In allowing scholars to identify actions, thoughts, and speech acts as distinct characteristics of international crises while also granting flexibility about the inclusion of a non-random sample of novel variables, our approach is a notable improvement over the state of the art.

@@

Our strategy for dealing with both problems is a plausibly objective yardstick called a synthetic historical narrative. We collect a large diverse corpus of narratives spanning timelines, encyclopedia entries, journal articles, news reports, websites, and government documents. Using natural language processing (fully described in SI Appendix 3.1), we identify details that appear across multiple accounts. A detail refers to the smallest textual unit (a word or n-gram) for which we can calculate similarity across corpora to identify whether sentences semantically refer to the same broader observed event [@narayanDontGiveMe2018]. The more accounts that mention a detail, the more central it is to understanding the true historical episode. The theoretical motivation is that authors face word limits which force them to pick and choose which details to include, and they choose details that serve the specific context of the document they are producing. With a sufficiently large and diverse corpus of documents, we can vary the context while holding the overall episode constant and see which details tend to be invariant to context. Sufficiently similar details were binned together and then summarized so they could be compared to the coding in ICBe. This presents a harder evaluation baseline than comparing ICBe's recall to just that of ICB since there are non-crisis aspects of these events that may be included in other narratives but are out of the scope of our data. For example, the nationalization of businesses in Cuba may be included as important context in the Cuban Missile Crisis in documents that do not focus on the crisis dimensions like ICB. Using this hard case, a recall measure of ICBe on the synthetic narratives thus serves as a way to evaluate the breadth of ICBe's ontology and potential application to non-crisis international events.

### 2023-11-20 replication_paper/psrm_draft/ICBe_manuscript.tex
AG rr 2 edits

\hhline{>{\arrayrulecolor[HTML]{FFFFFF}\global\arrayrulewidth=0pt}-~~|~~~~~~~|~~~}

@@

A unique feature of the ontology is that thought, speech, and do
behaviors can be nested into combinations, e.g.~an offer for the
U.S.S.R. to remove missiles from Cuba in exchange for the U.S. removing
missiles from Turkey. Through compounding, the ontology can capture what
players were said to have known, learned, or said about other specific
fully described actions. Two other novel features of our data -- its
re-programmamable coding and its inclusion of thought, speech, and
action -- intersect to provide a unique contribution. No existing event
data distinguishes thoughts, speeches, and actions and most are limited
to just coding actions given they are the most readily observable and
thus the full universe could at least theoretically be
included.\footnote{Even the coding of overt actions like MIDs is not
  without contention (Gibler 2018).} Yet, methods like archival work and
expert interviews continue to improve identification of thoughts and
speeches that help us understand the reasoning process underlying
political decisions and what actually occurred behind the scenes. This
is an improvement over the state of the art that omits these by design
in allowing scholars to identify thoughts associated with things like
crisis escalation as well as patterns of omission to be advanced in
future work, but the modular nature of our coding means the inclusion of
a non-random sample of thoughts and speeches is neither necessary nor
does it have inevitable downstream effects on the rest of the coding in
ways that produce unwanted bias.

@@

Our strategy for dealing with both problems is a plausibly objective
yardstick called a synthetic historical narrative. We collect a large
diverse corpus of narratives spanning timelines, encyclopedia entries,
journal articles, news reports, websites, and government documents.
Using natural language processing (fully described in SI Appendix 3.1),
we identify details that appear across multiple accounts. The more
accounts that mention a detail, the more central it is to understanding
the true historical episode. The theoretical motivation is that authors
face word limits which force them to pick and choose which details to
include, and they choose details that serve the specific context of the
document they are producing. With a sufficiently large and diverse
corpus of documents, we can vary the context while holding the overall
episode constant and see which details tend to be invariant to context.
This presents a harder evaluation baseline than comparing ICBe's recall
to just that of ICB since there are non-crisis aspects of these events
that may be included in other narratives but are out of the scope of our
data. For example, the nationalization of businesses in Cuba may be
included as important context in the Cuban Missile Crisis in documents
that do not focus on the crisis dimensions like ICB. Using this hard
case, a recall measure of ICBe on the synthetic narratives thus serves
as a way to evaluate the breadth of ICBe's ontology and potential
application to non-crisis international events.

@@

@@

>>>>

\hhline{>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=0pt}-~~|~~~~~~~|~~~}

@@

A unique feature of the ontology is that thought, speech, and do
behaviors can be nested into combinations, e.g.~an offer for the
U.S.S.R. to remove missiles from Cuba in exchange for the U.S. removing
missiles from Turkey. Through compounding, the ontology can capture what
players were said to have known, learned, or said about other specific
fully described actions. Two other novel features of our data -- its
re-programmamable coding and its inclusion of thought, speech, and
action -- intersect to provide a unique contribution.

First, re-programmable coding means scholars can make modifications to
the ontology we provide and re-run the code script to produce a modified
dataset. For example, if thoughts and speech acts are immaterial for a
particular research question, those higher-level categories can be
modified in a way that does not produce inevitable downstream effects on
the rest of the coding. This is necessary because unlike existing
datasets where variable values are largely independent, the nested
nature of the ICBe variables means filtering out speech acts also
requires filtered out variables about the actors, locations, and timing
of those speech acts.

Second, no existing event data distinguishes thoughts, speeches, and
actions. In fact, most only try to code actions and entirely omit
thoughts and speech acts despite recognition of their importance in
international politics (Smith 1998). Scholars have opted against coding
thoughts and speech acts because of a lack of confidence the full
universe could be readily observed and consequently at least
theoretically be included.\footnote{Even the coding of overt actions
  like MIDs is not without contention (Gibler 2018).} But the perfect
should not be the enemy of the good, and measurement challenges are only
overcome after an initial attempt to estimate difficult to observe
concepts of interest. The ICB narratives are one of the better sources
for this endeavor due to the consistent use of high-quality primary
source material that takes advantage of qualitative methods well-suited
to identifying thoughts and speech acts like archival work and expert
interviews. Important questions about potential bias remain like whether
a thought was known at the time by those involved or only recorded long
afterwards by historians, but data validation required there to be data
to be validated. By design, existing datasets largely omit thoughts and
speech acts and also treat the resulting dataset as the static product
of hardwired coding decisions. In allowing scholars to identify actions,
thoughts, and speech acts as distinct characteristics of international
crises while also granting flexibility about the inclusion of a
non-random sample of novel variables, our approach is a notable
improvement over the state of the art.

@@

Our strategy for dealing with both problems is a plausibly objective
yardstick called a synthetic historical narrative. We collect a large
diverse corpus of narratives spanning timelines, encyclopedia entries,
journal articles, news reports, websites, and government documents.
Using natural language processing (fully described in SI Appendix 3.1),
we identify details that appear across multiple accounts. A detail
refers to the smallest textual unit (a word or n-gram) for which we can
calculate similarity across corpora to identify whether sentences
semantically refer to the same broader observed event (Narayan, Cohen,
and Lapata 2018). The more accounts that mention a detail, the more
central it is to understanding the true historical episode. The
theoretical motivation is that authors face word limits which force them
to pick and choose which details to include, and they choose details
that serve the specific context of the document they are producing. With
a sufficiently large and diverse corpus of documents, we can vary the
context while holding the overall episode constant and see which details
tend to be invariant to context. Sufficiently similar details were
binned together and then summarized so they could be compared to the
coding in ICBe. This presents a harder evaluation baseline than
comparing ICBe's recall to just that of ICB since there are non-crisis
aspects of these events that may be included in other narratives but are
out of the scope of our data. For example, the nationalization of
businesses in Cuba may be included as important context in the Cuban
Missile Crisis in documents that do not focus on the crisis dimensions
like ICB. Using this hard case, a recall measure of ICBe on the
synthetic narratives thus serves as a way to evaluate the breadth of
ICBe's ontology and potential application to non-crisis international
events.

@@

\leavevmode\vadjust pre{\hypertarget{ref-narayanDontGiveMe2018}{}}%
Narayan, Shashi, Shay B Cohen, and Mirella Lapata. 2018. {``Don't Give
Me the Details, Just the Summary! Topic-Aware Convolutional Neural
Networks for Extreme Summarization.''} \emph{arXiv Preprint
arXiv:1808.08745}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-smithInternationalCrisesDomestic1998}{}}%
Smith, Alastair. 1998. {``International {Crises} and {Domestic
Politics}.''} \emph{American Political Science Review} 92 (3): 623--38.
\url{https://doi.org/10.2307/2585485}.
